---
title: "Beauty, or not, in the classroom"
author: "Varun Venkatesh"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Load Required Libraries
```{r}
library(ggplot2)
library(dplyr)
library(knitr)
```

## Load Data
```{r}
profs <- read.csv("profs.csv")
kable(head(profs))
```

## Part A: Histogram of Course Evaluation Scores
```{r}
ggplot(profs, aes(x = eval)) +
  geom_histogram(binwidth = 0.2, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Course Evaluation Scores", 
       x = "Evaluation Score", y = "Frequency") +
  theme_minimal()
```

*Caption:* This histogram displays the distribution of course evaluation scores ranging from 1 to 5, with the majority clustering in the mid-to-high range. This pattern suggests that most instructors receive good evaluations from students.

## Part B: Boxplots of Evaluation Scores by Native English Speaker Status
```{r}
ggplot(profs, aes(x = as.factor(native), y = eval, fill = as.factor(native))) +
  geom_boxplot() +
  labs(title = "Course Evaluation Scores by Native English Speaker Status", 
       x = "Native English Speaker", y = "Evaluation Score") +
  scale_x_discrete(labels = c("Non-Native", "Native")) +
  theme_minimal() +
  scale_fill_manual(values = c("red", "green"))
```

*Caption:* The boxplot compares evaluation scores between native and non-native English-speaking professors. Native speakers tend to receive slightly higher median scores, and their ratings show less variability, showing biases in student perceptions.

## Part C: Faceted Histogram of Evaluation Scores by Gender
```{r}
ggplot(profs, aes(x = eval)) +
  geom_histogram(binwidth = 0.2, fill = "purple", color = "black", alpha = 0.7) +
  facet_wrap(~ gender, nrow = 2) +
  labs(title = "Distribution of Course Evaluation Scores by Gender", 
       x = "Evaluation Score", y = "Frequency") +
  theme_minimal()
```

*Caption:* The faceted histogram shows the distribution of evaluation scores for male and female instructors. Both genders have similar rating patterns, but there are small differences in density.


## Part D: Scatterplot of Attractiveness vs. Evaluation Scores
```{r}
ggplot(profs, aes(x = beauty, y = eval)) +
  geom_point(alpha = 0.5, color = "blue") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Relationship Between Physical Attractiveness and Course Evaluation Scores", 
       x = "Attractiveness Score", y = "Evaluation Score") +
  theme_minimal()
```

*Caption:* This scatterplot shows the relationship between instructors' physical attractiveness (x-axis) and their evaluation scores (y-axis). A positive trend line indicates a potential association, suggesting that perceived attractiveness might influence evaluation scores.

