---
title: "Capital Metro UT Ridership Analysis"
author: "Varun Venkatesh"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Load Required Libraries
```{r}
library(ggplot2)
library(dplyr)
library(knitr)
```


## Load Data
```{r}
capmetro_UT <- read.csv("capmetro_UT.csv")

# Reorder categorical variables
capmetro_UT <- mutate(capmetro_UT,
  day_of_week = factor(day_of_week, levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")),
  month = factor(month, levels = c("Sep", "Oct", "Nov"))
)
kable(head(capmetro_UT))
```

## Plot 1: Faceted Line Graph of Average Boardings by Hour, Day, and Month
```{r}
avg_boardings <- capmetro_UT %>%
  group_by(hour_of_day, day_of_week, month) %>%
  summarise(avg_boarding = mean(boarding), .groups = "drop")

ggplot(avg_boardings, aes(x = hour_of_day, y = avg_boarding, color = month)) +
  geom_line(size = 1) +
  facet_wrap(~ day_of_week, ncol = 2) +
  labs(title = "Average Boardings by Hour, Day of Week, and Month",
       x = "Hour of the Day", y = "Average Boardings",
       color = "Month") +
  scale_x_continuous(breaks = seq(6, 22, by = 2)) +
  theme_minimal()
```

*Caption:* This faceted line graph shows average bus boardings by hour of the day, day of the week, and month. Peak boarding times remain consistent at 8-9 AM and 5-6 PM across most days. Lower boardings on Mondays in September are probably because of holidays or semester start days, while reductions on Wednesdays to Fridays in November are probably due to Thanksgiving break.

## Plot 2: Faceted Scatter Plot of Boardings vs. Temperature
```{r}
ggplot(capmetro_UT, aes(x = temperature, y = boarding, color = as.factor(weekend))) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ hour_of_day, ncol = 4) +
  labs(title = "Boardings vs. Temperature by Hour of Day",
       x = "Temperature (°F)", y = "Boardings",
       color = "Day Type") +
  scale_color_manual(values = c("blue", "red"), labels = c("Weekday", "Weekend")) +
  theme_minimal()
```

*Caption:* This faceted scatterplot examines the influence of temperature (x-axis) on bus boardings (y-axis) across different hours of the day. Colder temperatures reduce ridership, especially during early mornings and late evenings. However, ridership during midday hours remains relatively stable, probably due to students and 9 to 5 commuters.
---

