---
title: "Bike Sharing Analysis"
author: "Varun Venkatesh"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Required Libraries
```{r}
library(ggplot2)
library(dplyr)
library(knitr)
```

## Load Data
```{r}
bikeshare <- read.csv("bikeshare.csv")
kable(head(bikeshare))
```

## Plot A: Line Graph of Average Hourly Bike Rentals
```{r}
avg_hourly_rentals <- bikeshare %>%
  group_by(hr) %>%
  summarise(avg_total = mean(total))

ggplot(avg_hourly_rentals, aes(x = hr, y = avg_total)) +
  geom_line(color = "blue", size = 1) +
  labs(title = "Average Hourly Bike Rentals",
       x = "Hour of the Day", y = "Average Rentals") +
  theme_minimal()
```

*Caption:* This line graph illustrates the average number of bike rentals for each hour of the day. Ridership peaks during morning (7-9 AM) and evening (5-7 PM) commute hours, showing that it is used heavily by commuters.

## Plot B: Faceted Line Graph of Bike Rentals by Hour and Working Day
```{r}
avg_workingday_rentals <- bikeshare %>%
  group_by(hr, workingday) %>%
  summarise(avg_total = mean(total))

ggplot(avg_workingday_rentals, aes(x = hr, y = avg_total, color = as.factor(workingday))) +
  geom_line(size = 1) +
  facet_wrap(~ workingday, nrow = 2) +
  labs(title = "Average Bike Rentals by Hour (Working vs. Non-Working Day)",
       x = "Hour of the Day", y = "Average Rentals") +
  theme_minimal()
```

*Caption:* The faceted line graph shows hourly bike rental patterns on working and non-working days. On working days, rentals peak during commuting hours, but non-working days have more evenly distributed activity throughout the day, showing that the use is more for leisure than for work.

## Plot C: Faceted Bar Plot of 9 AM Ridership by Weather Situation
```{r}
ridership_9am <- bikeshare %>%
  filter(hr == 9) %>%
  group_by(weathersit, workingday) %>%
  summarise(avg_total = mean(total))

ggplot(ridership_9am, aes(x = as.factor(weathersit), y = avg_total, fill = as.factor(weathersit))) +
  geom_bar(stat = "identity") +
  facet_wrap(~ workingday, nrow = 2) +
  labs(title = "9 AM Ridership by Weather Situation",
       x = "Weather Condition", y = "Average Rentals") +
  scale_x_discrete(labels = c("Clear", "Mist", "Light Snow/Rain", "Heavy Rain/Snow")) +
  theme_minimal()
```

*Caption:* This faceted bar plot shows the average 9 AM ridership under different weather conditions, separated by working and non-working days. Clear weather is the highest ridership, while cold/wet conditions like mist or rain significantly reduce usage, especially on non-working days.
